Kronos \& Krios: Evidence for accretion of a massive, rocky planetary system in a comoving pair of solar-type stars

### 2017-09-12 paper/ms.tex
minor edits per JMB comments

Binary stars that form from the same birth cloud start with nearly identical
composition.
A differential analysis of the chemical composition of binary stars can reveal
their history through the chemical signatures related to planet formation or
accretion regardless of Galactic chemical evolution.
\askjmb{
For example, \citealt{Fischer:2005aa} showed that higher metallicity
stars more likely host close-in giant planets, which likely promotes
the formation of giant planets by increasing the availability of small particle
condensates that form planetesimals.
}
On the other hand, if host stars are polluted after their birth by rocky
planetary material with high refractory-to-volatile ratio, the convective
envelope of the stars may be enhanced in refractory elements (e.g., \elem{Fe})
compared to their initial state (e.g., \citealt{Pinsonneault:2001aa}).
Thus, differences in planet formation or accretion in two otherwise identical
stars may imprint differences in chemical abundances that depend on the
condensation temperature (\Tcondens).

@@

\begin{figure}[htpb]
  \centering
  \includegraphics[width=0.95\linewidth]{deltaXH_elem_violins.pdf}
  \caption{Abundance difference in this pair and other twin-like
    ($\Delta T_\mathrm{eff}\lesssim 100$~K) wide binaries in
    \citealt{2016ApJS..225...32B}.
    The differences in other pairs are small ($<0.05$~dex)
    for all elements except \elem{N} and \elem{O} which are the most
    uncertain, making the difference of $\approx 0.2$~dex seen in
    \bizarreone-\sunanalog\ rare.
    Additionally, we show the distribution of abundance differences
    between field stars with similar metallicity difference
    ($\Delta[\elem{Fe}/\elem{H}] \approx 0.2$)
    as violins with medians indicated by black line segments.
    These are random pairings of single stars in
    in \citealt{2016ApJS..225...32B} at two metallicity bins,
    $-0.025 < \feh < 0.025$ (160 stars) and $0.175 > \feh > 0.225$ (137 stars),
    similar to \bizarreone\ and \sunanalog.
    The difference is always taken to be
    $\mathrm{higher}\,\feh - \mathrm{lower}\,\feh$.
    Thus, the narrower range of in $\Delta\feh$ is by construction.
    Random pairings of disk stars with similar $\Delta\feh$ usually show
    similar enhancement in all other elements
    unlike the pattern seen in \bizarreone-\sunanalog\ pair.
  }
  \label{fig:deltaXH}
\end{figure}

@@

{\bf
The surface lithium abundance in a Sun-like star decreases with its age due to
mixing induced by convection or rotation, which brings the lithium into the
interior ($T>2.5 \times 10^{6}$~K) where it will be destroyed by proton capture
burning.
Thus, surface lithium abundance can be an indicator of stellar ages,
especially whether the star is very young ($\lesssim 1$~Gyr).
The absolute $\elem{Li}$ abundance of solar-type stars also correlates steeply
with the effective temperature (e.g., \citealt{Chen2001,2012ApJ...756...46R}).
Generally, cooler stars with larger convective envelope have lower \elem{Li}
abundances.
The absolute $\elem{Li}$ abundance of $2.25$~dex for \sunanalog\
is typical for its \teff.
On the other hand, the lithium abundance ($A(\elem{Li}) = 2.75$) of \bizarreone,
which has lower \teff\ than \sunanalog, is not only higher than that of
\sunanalog\ but also much higher compared to other field stars of similar \teff.
Given the overall higher metal abundances and the peculiar abundance patterns
in \bizarreone, it is unclear, however, whether this higher $\elem{Li}$
abundance means a younger age or something else.
For example, \citealt{Casey:2016aa} attributes the presence of $\elem{Li}$-rich
red giant stars to the engulfment of substellar companions such as gas giant
planets or brown dwarfs which may replenish $\elem{Li}$.
}

@@

In this subsection, we explore the hypothesis that chemical inhomogeneity
within the birth cloud is the source of the observed abundance difference.
There is ample evidence against this scenario as most wide binaries show in
\feh\ less than $0.02$~dex (\citealt{Desidera:2004aa,Gratton:2001aa}.
In the cases in which a significant difference is detected
with high-precision abundance measurements (see \sectionname~\ref{app:review}),
the difference is typically $\sim 0.05$~dex.
Consistent with these results,
none of the other seven similar wide binaries examined in
\citealt{2016ApJS..225...32B} show the same level of differences in abundances.
Though there is generally a larger spread in $\elem{C}$, $\elem{N}$ and
$\elem{O}$, and some pairs show a difference in particular elements as large as
$\approx 0.15$~dex.
The median and maximum \feh\ difference between component stars in the other
seven pairs is $0.02$~dex and $0.09$~dex, respectively.
The differences are even smaller (maximum $\Delta\feh = 0.03$~dex) if we
compare only twin-like ($\Delta T_\mathrm{eff} \lesssim 100$~K) pairs
(\figname~\ref{fig:deltaXH}, black lines).
Thus, a difference of $\approx 0.2$~dex seen in \bizarreone-\sunanalog\ pair is
unlikely to be due to chemical inhomogeneity in the birth cloud.

@@

What about \elem{Li}?
The element \elem{Li} is worth special attention in the context of the
accretion scenario.
Because Li is present in either carbonaceous chondrites or bulk Earth with a
concentration of $1-1.5$~ppm in mass (\citealt{2003TrGeo...2..547M}), but
is depleted quickly within a few gigayears on the surface of a
Sun-like star (\askjmb{CITE?}), accretion of either material at later times
will significantly replenish the lithium on the star's surface.
For the present-day Sun ($A(\elem{Li}) = 1.05$), the accretion of \maccreted\ of
bulk Earth-like material would result in $\Delta\elemH{Li} \approx 1.65$~dex
(see the inset of \figname~\ref{fig:toycalc}).
This closely matches what we find: the \elem{Li} abundance of \bizarreone\ is
$A(\elem{Li}) = 2.75$ (Table~\ref{tab:kk}, \citealt{jmlithium})
approximately $1.7$~dex higher than the solar value.

@@

{\bf
Finally, we mention that if the hypothesized accretion is
\askjmb{recent (HOW RECENT?)},
detection of $^6 \elem{Li}$ provides a strong test for this
scenario.
This isotope of \elem{Li} is destroyed at even lower temperatures than
$^7\elem{Li}$, and theoretically expected to be absent
(\citealt{1997ARA&A..35..557P}).
However, an accretion of rocky material could have replenished $^6 \elem{Li}$.
Because $^6\elem{Li}$ lines are slightly longer in wavelengths,
presence of $^6\elem{Li}$ increases the asymmetry of \elem{Li} $6707.6$ feature.
This is a very subtle effect that requires a careful modelling effort
and a higher signal-to-noise, higher resolution spectra
(see e.g., \citealt{Israelian:2001,2002MNRAS.335.1005R}).
Such investigation was not warranted by the current data (\citealt{jmlithium}).
}

@@

The two stars have not been included in any publicly released data from planet
search programs.
If both stars have accreted planetary material, it would be very interesting to
search for the existence and architectures of the planetary systems left
behind.

>>>>

Binary stars that form from the same birth cloud start with nearly identical
composition.
A differential analysis of the chemical composition of binary stars can reveal
their history through the chemical signatures related to planet formation or
accretion regardless of Galactic chemical evolution.
Stars hosting giant planets on short period orbits have been shown, via
population studies, to form more readily around inherently metal rich stars
(\citealt{Fischer:2005aa,Santos2004}).
However, the post-formation accretion of rocky planets can still alter the
photospheric abundances.
If host stars are polluted after their birth by rocky planetary material with a
high refractory-to-volatile ratio, the convective envelope of the stars may be
enhanced in refractory elements (e.g., \elem{Fe}) compared to their initial
state (e.g., \citealt{Pinsonneault:2001aa}).
Thus, differences in planet formation or accretion in two otherwise identical
stars may imprint differences in chemical abundances that depend on the
condensation temperature (\Tcondens).

@@

\begin{figure}[htpb]
  \centering
  \includegraphics[width=0.95\linewidth]{deltaXH_elem_violins.pdf}
  \caption{Abundance difference in this pair and other twin-like
    ($\Delta T_\mathrm{eff}\lesssim 100$~K) wide binaries in
    \citealt{2016ApJS..225...32B}.
    The differences in other pairs are small ($<0.05$~dex)
    for all elements except \elem{N} and \elem{O} which are the most
    uncertain, making the difference of $\approx 0.2$~dex seen in
    \bizarreone-\sunanalog\ rare.
    Additionally, we show the distribution of abundance differences
    between field stars with similar metallicity difference
    ($\Delta[\elem{Fe}/\elem{H}] \approx 0.2$)
    as violins with medians indicated by black line segments.
    These are random pairings of single stars in
    in \citealt{2016ApJS..225...32B} at two metallicity bins,
    $-0.025 < \feh < 0.025$ (160 stars) and $0.175 > \feh > 0.225$ (137 stars),
    similar to \bizarreone\ and \sunanalog.
    The difference is always taken to be
    $\mathrm{higher}\,\feh - \mathrm{lower}\,\feh$.
    Thus, the narrower range of $\Delta\feh$ is by construction.
    Random pairings of disk stars with similar $\Delta\feh$ usually show
    similar enhancement in all other elements
    unlike the pattern seen in \bizarreone-\sunanalog\ pair.
  }
  \label{fig:deltaXH}
\end{figure}

@@

The surface lithium abundance in a Sun-like star decreases with its age due to
mixing induced by convection or rotation, which brings the lithium into the
interior ($T>2.5 \times 10^{6}$~K) where it will be destroyed by proton capture
burning.
In hotter stars with thin convective zones on the main sequence, most of this
mixing occurs in the pre-main sequence phase when the star is fully convective.
% TODO: REF?
Thus, surface lithium abundance can be an indicator of stellar ages,
especially whether the star is very young ($\lesssim 1$~Gyr).
The absolute $\elem{Li}$ abundance of solar-type stars also correlates steeply
with the effective temperature (e.g., \citealt{Chen2001,2012ApJ...756...46R}).
Generally, cooler stars with larger convective envelope have lower \elem{Li}
abundances.
The absolute $\elem{Li}$ abundance of $2.25$~dex for \sunanalog\
is typical for its \teff.
On the other hand, the lithium abundance ($A(\elem{Li}) = 2.75$) of \bizarreone,
which has lower \teff\ than \sunanalog, is not only higher than that of
\sunanalog\ but also much higher compared to other field stars of similar \teff.
Given the overall higher metal abundances and the peculiar abundance patterns
in \bizarreone, it is unclear, however, whether this higher $\elem{Li}$
abundance means a younger age or something else.
For example, \citealt{Casey:2016aa} attributes the presence of $\elem{Li}$-rich
red giant stars to the engulfment of substellar companions such as gas giant
planets or brown dwarfs which may replenish $\elem{Li}$.

@@

In this subsection, we explore the hypothesis that chemical inhomogeneity
within the birth cloud is the source of the observed abundance difference.
There is ample evidence against this scenario as most wide binaries show a
difference in \feh\ less than $0.02$~dex
(\citealt{Desidera:2004aa,Gratton:2001aa}.
In the cases in which a significant difference is detected
with high-precision abundance measurements (see \sectionname~\ref{app:review}),
the difference is typically $\sim 0.05$~dex.
Consistent with these results,
none of the other seven similar wide binaries examined in
\citealt{2016ApJS..225...32B} show the same level of differences in abundances.
Though there is generally a larger spread in $\elem{C}$, $\elem{N}$ and
$\elem{O}$, and some pairs show a difference in particular elements as large as
$\approx 0.15$~dex.
The median and maximum \feh\ difference between component stars in the other
seven pairs is $0.02$~dex and $0.09$~dex, respectively.
The differences are even smaller (maximum $\Delta\feh = 0.03$~dex) if we
compare only twin-like ($\Delta T_\mathrm{eff} \lesssim 100$~K) pairs
(\figname~\ref{fig:deltaXH}, black lines).
Thus, a difference of $\approx 0.2$~dex seen in \bizarreone-\sunanalog\ pair is
unlikely to be due to chemical inhomogeneity in the birth cloud.

@@

What about \elem{Li}?
The element \elem{Li} is worth special attention in the context of the
accretion scenario.
Because Li is present in either carbonaceous chondrites or bulk Earth with a
concentration of $1-1.5$~ppm in mass (\citealt{2003TrGeo...2..547M}), but is
depleted quickly within the first Gyr on the surface of a Sun-like star
(\citealt{Thevenin2017,Baraffe2017}), accretion of either material at later
times will significantly replenish the lithium on the star's surface.
For the present-day Sun ($A(\elem{Li}) = 1.05$), the accretion of \maccreted\ of
bulk Earth-like material would result in $\Delta\elemH{Li} \approx 1.65$~dex
(see the inset of \figname~\ref{fig:toycalc}).
This closely matches what we find: the \elem{Li} abundance of \bizarreone\ is
$A(\elem{Li}) = 2.75$ (Table~\ref{tab:kk}, \citealt{jmlithium})
approximately $1.7$~dex higher than the solar value.

@@

Finally, we mention that detection of $^6 \elem{Li}$ provides a strong test for
this scenario.
This isotope of \elem{Li} is destroyed at even lower temperatures than
$^7\elem{Li}$, and theoretically expected to be absent
(\citealt{1997ARA&A..35..557P}).
However, an accretion of rocky material could have replenished $^6 \elem{Li}$.
Because $^6\elem{Li}$ lines are slightly longer in wavelengths,
presence of $^6\elem{Li}$ increases the asymmetry of \elem{Li} $6707.6$ feature.
Depending on how recent the accretion was and how fast $^6 \elem{Li}$ is
depleted on the main sequence, this feature may be detectable.
This is a very subtle effect that requires a higher signal-to-noise, higher
resolution spectra and careful modelling effort
(see e.g., \citealt{Israelian:2001,2002MNRAS.335.1005R}).
Such investigation was not warranted by the current data (\citealt{jmlithium}).

@@

The two stars have not been included in any publicly released data from planet
search programs.
We have begun a precision radial velocity campaign for the two stars and early
indications are that there are no close in giant planets.
If both stars have accreted planetary material, it would be very interesting to
search for the existence and architectures of the planetary systems left
behind.
